Sorting by Reversals

### 2020-07-03 slides.tex
update examples

\usepackage[sorting=ynt]{biblatex}
\addbibresource{sources.bib}

@@

\section{Introduction}

@@

\begin{exampleblock}{Example}
Let $\pi = (0\ 4\ 2\ 1\ 3\ 5) \in S_4$. \par
Then
\begin{align*}
    \pi \circ \rho(1,4) = (0\ \r{3\ 1\ 2\ 4}\ 5).
\end{align*}
\end{exampleblock}

@@

\begin{exampleblock}{Example}
Let $\pi = (0\ 4\ 2\ 1\ 3\ 5) \in S_4$.\pause
\begin{align*}
    \pi \circ \rho(1,3) &= (0\ \r{1\ 2\ 4}\ 3\ 5) \\ \pause
    \pi \circ \rho(1,3) \circ \rho(3,4) &= (0\ 1\ 2\ \r{3\ 4}\ 5) = id
\end{align*}\pause
$\implies d(\pi) \leq 2$.
\end{exampleblock}

@@

\begin{exampleblock}{Example}
Let $\pi = (0\ \r\mid\ 4\ \r\mid\ 2\ 1\ \r\mid\ 3\ \r\mid\ 5) \in S_4$. \pause Then $G(\pi)$ is

\begin{center}
\begin{tikzpicture}
	\begin{pgfonlayer}{nodelayer}
		\node [style=node] (0) at (-3, 0) {0};
		\node [style=node] (3) at (-1.5, 0) {4};
		\node [style=node] (4) at (0, 0) {2};
		\node [style=node] (5) at (1.5, 0) {1};
		\node [style=node] (8) at (3, 0) {3};
		\node [style=node] (9) at (4.5, 0) {5};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=red edge] (3) to (4);
		\draw [style=red edge] (8) to (9);
		\draw [style=red edge] (0) to (3);
		\draw [style=red edge] (5) to (8);
		\draw [style=blue edge, bend left] (4) to (8);
		\draw [style=blue edge, bend left] (3) to (9);
		\draw [style=blue edge, bend right] (0) to (5);
		\draw [style=blue edge, bend right] (3) to (8);
	\end{pgfonlayer}
\end{tikzpicture}
\end{center}
\end{exampleblock}\pause

@@

\begin{center}
\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=node] (0) at (-3, 0) {0};
		\node [style=node] (3) at (-1.5, 0) {4};
		\node [style=node] (4) at (0, 0) {2};
		\node [style=node] (5) at (1.5, 0) {1};
		\node [style=node] (8) at (3, 0) {3};
		\node [style=node] (9) at (4.5, 0) {5};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=red edge] (3) to (4);
		\draw [style=red edge] (8) to (9);
		\draw [style=red edge] (0) to (3);
		\draw [style=red edge] (5) to (8);
		\draw [style=blue edge, bend left] (4) to (8);
		\draw [style=blue edge, bend left] (3) to (9);
		\draw [style=blue edge, bend right] (0) to (5);
		\draw [style=blue edge, bend right] (3) to (8);
	\end{pgfonlayer}
\end{tikzpicture}
\end{center}

@@

\begin{definition}[reversal graph, \citeauthor*{Christie1998}]
Given an alternating cycle decomposition $C$ of $G(\pi)$, \par\pause
define $R(C)$ with
\begin{itemize}
    \item an isolated \b{blue} vertex for each adjacency in $\pi$\pause;
    \item $m$ vertices for each $m$-cycle in $C$ each representing the reversal $\rho(u)$ acting on the two red edges connected by a blue edge\pause;
    \begin{itemize}
        \item a vertex is \r{red} if the represented reversal is a $1$- or $2$-reversal
        \item a vertex is \b{blue} otherwise
    \end{itemize}\pause
    \item connect two vertices with an edge if their corresponding blue edges \textit{interleave} (i.e. overlap).
\end{itemize}
\end{definition}\pause

\begin{corollary}[\citeauthor*{Christie1998}]
The only alternating cycle decomposition of $id$ is $C = \emptyset$. \par
$R(C)$ consists of $n+1$ isolated blue vertices.
\end{corollary}

@@

\begin{tabularx}{\textwidth}{XR}
{\footnotesize\begin{align*}
    C = \{&\{\{0,4\},\{3,4\},\{1,3\},\{0,1\}\} \\
          &\{\{2,4\},\{2,3\},\{3,5\},\{4,5\}\}\}
\end{align*}} &
\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=node] (0) at (-3, 0) {0};
		\node [style=node] (3) at (-1.5, 0) {4};
		\node [style=node] (4) at (0, 0) {2};
		\node [style=node] (5) at (1.5, 0) {1};
		\node [style=node] (8) at (3, 0) {3};
		\node [style=node] (9) at (4.5, 0) {5};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=red edge] (3) to (4);
		\draw [style=red edge] (8) to (9);
		\draw [style=red edge] (0) to (3);
		\draw [style=red edge] (5) to (8);
		\draw [style=blue edge, bend left] (4) to (8);
		\draw [style=blue edge, bend left] (3) to (9);
		\draw [style=blue edge, bend right] (0) to (5);
		\draw [style=blue edge, bend right] (3) to (8);
	\end{pgfonlayer}
\end{tikzpicture}
\end{tabularx}\pause

@@

\begin{center}
\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=red node] (0) at (0, 1) {$\{3,4\}$};
		\node [style=red node] (1) at (-2, 1) {$\{0,1\}$};
		\node [style=blue node] (2) at (-2, -1) {$\{2,3\}$};
		\node [style=blue node] (3) at (0, -1) {$\{4,5\}$};
		\node [style=blue node] (4) at (2, 0) {$\{1,2\}$};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw (1) to (0);
		\draw (2) to (1);
		\draw (0) to (3);
		\draw (1) to (3);
		\draw (2) to (0);
	\end{pgfonlayer}
\end{tikzpicture}
\end{center}
\end{exampleblock}\pause

@@

\begin{exampleblock}{Example}
Let $\pi = (0\ 4\ 2\ 1\ 3\ 5) \in S_4$ and $u = \{0,1\}$.\padding

\begin{tabularx}{\textwidth}{X|R}
$G(\pi)$ and $R(C)$ & $G(\pi \circ \rho(u))$ and $R(C_u)$ \\ \hline
\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=node] (0) at (-3, 0) {0};
		\node [style=node] (3) at (-1.5, 0) {4};
		\node [style=node] (4) at (0, 0) {2};
		\node [style=node] (5) at (1.5, 0) {1};
		\node [style=node] (8) at (3, 0) {3};
		\node [style=node] (9) at (4.5, 0) {5};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=red edge] (3) to (4);
		\draw [style=red edge] (8) to (9);
		\draw [style=red edge] (0) to (3);
		\draw [style=red edge] (5) to (8);
		\draw [style=blue edge, bend left] (4) to (8);
		\draw [style=blue edge, bend left] (3) to (9);
		\draw [style=blue edge, bend right] (0) to (5);
		\draw [style=blue edge, bend right] (3) to (8);
	\end{pgfonlayer}
\end{tikzpicture} &
\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=node] (0) at (-3, 0) {0};
		\node [style=node] (3) at (-1.5, 0) {1};
		\node [style=node] (4) at (0, 0) {2};
		\node [style=node] (5) at (1.5, 0) {4};
		\node [style=node] (8) at (3, 0) {3};
		\node [style=node] (9) at (4.5, 0) {5};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=red edge] (4) to (5);
		\draw [style=red edge] (8) to (9);
		\draw [style=blue edge, bend left] (4) to (8);
		\draw [style=blue edge, bend left] (5) to (9);
	\end{pgfonlayer}
\end{tikzpicture}\\
\spadding\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=red node] (0) at (0, 1) {$\{3,4\}$};
		\node [style=red node] (1) at (-2, 1) {$\{0,1\}$};
		\node [style=blue node] (2) at (-2, -1) {$\{2,3\}$};
		\node [style=blue node] (3) at (0, -1) {$\{4,5\}$};
		\node [style=blue node] (4) at (2, 0) {$\{1,2\}$};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw (1) to (0);
		\draw (2) to (1);
		\draw (0) to (3);
		\draw (1) to (3);
		\draw (2) to (0);
	\end{pgfonlayer}
\end{tikzpicture} &
\spadding\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=blue node] (0) at (0, 1) {$\{3,4\}$};
		\node [style=blue node] (1) at (-2, 1) {$\{0,1\}$};
		\node [style=red node] (2) at (-2, -1) {$\{2,3\}$};
		\node [style=red node] (3) at (0, -1) {$\{4,5\}$};
		\node [style=blue node] (4) at (2, 0) {$\{1,2\}$};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw (2) to (3);
	\end{pgfonlayer}
\end{tikzpicture}
\end{tabularx}
\end{exampleblock}

>>>>

\usepackage[sorting=ynt,style=alphabetic]{biblatex}
\addbibresource{sources.bib}

@@

\section{Motivation}

@@

\begin{exampleblock}{Example}
Let $\pi = (0\ 1\ 3\ 4\ 2\ 5) \in S_4$. \par
Then
\begin{align*}
    \pi \circ \rho(2,4) = (0\ 1\ \r{2\ 4\ 3}\ 5).
\end{align*}
\end{exampleblock}

@@

\begin{exampleblock}{Example}
Let $\pi = (0\ 1\ 3\ 4\ 2\ 5) \in S_4$.\pause
\begin{align*}
    \pi \circ \rho(2,4) &= (0\ 1\ \r{2\ 4\ 3}\ 5) \\ \pause
    \pi \circ \rho(2,4) \circ \rho(3,4) &= (0\ 1\ 2\ \r{3\ 4}\ 5) = id
\end{align*}\pause
$\implies d(\pi) \leq 2$.
\end{exampleblock}

@@

\begin{exampleblock}{Example}
Let $\pi = (0\ 1\ \r\mid\ 3\ 4\ \r\mid\ 2\ \r\mid\ 5) \in S_4$. \pause Then $G(\pi)$ is

\spadding\begin{center}
\begin{tikzpicture}
	\begin{pgfonlayer}{nodelayer}
		\node [style=node] (0) at (-3, 0) {0};
		\node [style=node] (1) at (-1.5, 0) {1};
		\node [style=node] (2) at (0, 0) {3};
		\node [style=node] (3) at (1.5, 0) {4};
		\node [style=node] (4) at (3, 0) {2};
		\node [style=node] (5) at (4.5, 0) {5};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=red edge] (1) to (2);
		\draw [style=red edge] (3) to (4);
		\draw [style=red edge] (4) to (5);
		\draw [style=blue edge, bend right] (1) to (4);
		\draw [style=blue edge, bend left] (2) to (4);
		\draw [style=blue edge, bend left] (3) to (5);
	\end{pgfonlayer}
\end{tikzpicture}
\end{center}
\end{exampleblock}\pause

@@

\begin{center}
\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=node] (0) at (-3, 0) {0};
		\node [style=node] (1) at (-1.5, 0) {1};
		\node [style=node] (2) at (0, 0) {3};
		\node [style=node] (3) at (1.5, 0) {4};
		\node [style=node] (4) at (3, 0) {2};
		\node [style=node] (5) at (4.5, 0) {5};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=red edge] (1) to (2);
		\draw [style=red edge] (3) to (4);
		\draw [style=red edge] (4) to (5);
		\draw [style=blue edge, bend right] (1) to (4);
		\draw [style=blue edge, bend left] (2) to (4);
		\draw [style=blue edge, bend left] (3) to (5);
	\end{pgfonlayer}
\end{tikzpicture}
\end{center}

@@

\begin{definition}[reversal graph, \citeauthor*{Christie1998}]
Given an alternating cycle decomposition $C$ of $G(\pi)$, \par\pause
define $R(C)$ with
\begin{itemize}
    \item an isolated \b{blue} vertex for each adjacency in $\pi$\pause;
    \item $m$ vertices for each $m$-cycle in $C$ each representing the reversal $\rho(u)$ acting on the two red edges connected by a blue edge\pause;
    \begin{itemize}
        \item a vertex is \r{red} if the represented reversal is a $1$- or $2$-reversal
        \item a vertex is \b{blue} otherwise
    \end{itemize}\pause
    \item connect two vertices with an edge if their corresponding blue edges \textit{interleave}.
\end{itemize}
\end{definition}\pause

\begin{corollary}[\citeauthor*{Christie1998}]
The only alternating cycle decomposition of $G(id)$ is $C = \emptyset$. \par
$R(C)$ consists of $n+1$ isolated blue vertices.
\end{corollary}

@@

\begin{tabularx}{\textwidth}{XR}
{\footnotesize\begin{align*}
    C = \{(&\{1,3\},\{2,3\},\{2,4\}, \\
           &\{4,5\},\{2,5\},\{1,2\})\}
\end{align*}} &
\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=node] (0) at (-3, 0) {0};
		\node [style=node] (1) at (-1.5, 0) {1};
		\node [style=node] (2) at (0, 0) {3};
		\node [style=node] (3) at (1.5, 0) {4};
		\node [style=node] (4) at (3, 0) {2};
		\node [style=node] (5) at (4.5, 0) {5};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=red edge] (1) to (2);
		\draw [style=red edge] (3) to (4);
		\draw [style=red edge] (4) to (5);
		\draw [style=blue edge, bend right] (1) to (4);
		\draw [style=blue edge, bend left] (2) to (4);
		\draw [style=blue edge, bend left] (3) to (5);
	\end{pgfonlayer}
\end{tikzpicture}
\end{tabularx}\pause

@@

\begin{center}
\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=red node] (0) at (0, 1) {$\{2,3\}$};
		\node [style=red node] (1) at (-2, 1) {$\{1,2\}$};
		\node [style=blue node] (2) at (-1, -1) {$\{4,5\}$};
		\node [style=blue node] (3) at (2, -1) {$\{3,4\}$};
		\node [style=blue node] (4) at (2, 1) {$\{0,1\}$};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw (0) to (2);
		\draw (1) to (2);
	\end{pgfonlayer}
\end{tikzpicture}
\end{center}
\end{exampleblock}\pause

@@

\begin{exampleblock}{Example}
Let $\pi = (0\ 1\ 3\ 4\ 2\ 5) \in S_4$ and $u = \{1,2\}$.\padding

\begin{tabularx}{\textwidth}{X|R}
$G(\pi)$ and $R(C)$ & $G(\pi \circ \rho(u))$ and $R(C_u)$ \\ \hline
\spadding\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=node] (0) at (-3, 0) {0};
		\node [style=node] (1) at (-1.5, 0) {1};
		\node [style=node] (2) at (0, 0) {3};
		\node [style=node] (3) at (1.5, 0) {4};
		\node [style=node] (4) at (3, 0) {2};
		\node [style=node] (5) at (4.5, 0) {5};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=red edge] (1) to (2);
		\draw [style=red edge] (3) to (4);
		\draw [style=red edge] (4) to (5);
		\draw [style=blue edge, bend right] (1) to (4);
		\draw [style=blue edge, bend left] (2) to (4);
		\draw [style=blue edge, bend left] (3) to (5);
	\end{pgfonlayer}
\end{tikzpicture} &
\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=node] (0) at (-3, 0) {0};
		\node [style=node] (1) at (-1.5, 0) {1};
		\node [style=node] (2) at (0, 0) {2};
		\node [style=node] (3) at (1.5, 0) {4};
		\node [style=node] (4) at (3, 0) {3};
		\node [style=node] (5) at (4.5, 0) {5};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=red edge] (2) to (3);
		\draw [style=red edge] (4) to (5);
		\draw [style=blue edge, bend left] (2) to (4);
		\draw [style=blue edge, bend left] (3) to (5);
	\end{pgfonlayer}
\end{tikzpicture}\\
\spadding\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=red node] (0) at (0, 1) {$\{2,3\}$};
		\node [style=red node] (1) at (-2, 1) {$\{1,2\}$};
		\node [style=blue node] (2) at (-1, -1) {$\{4,5\}$};
		\node [style=blue node] (3) at (2, -1) {$\{3,4\}$};
		\node [style=blue node] (4) at (2, 1) {$\{0,1\}$};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw (0) to (2);
		\draw (1) to (2);
	\end{pgfonlayer}
\end{tikzpicture} &
\spadding\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=red node] (0) at (0, 1) {$\{2,3\}$};
		\node [style=blue node] (1) at (-2, 1) {$\{1,2\}$};
		\node [style=red node] (2) at (-1, -1) {$\{4,5\}$};
		\node [style=blue node] (3) at (2, -1) {$\{3,4\}$};
		\node [style=blue node] (4) at (2, 1) {$\{0,1\}$};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw (0) to (2);
	\end{pgfonlayer}
\end{tikzpicture}
\end{tabularx}
\end{exampleblock}
